risk-perception

### 2019-10-24 risk-percption.Rmd
simulation with mixed strategies

Given these payoffs, what's the point of indifference between the two crop types?

```{r, echo = FALSE}
plt_dat <- tibble(risk = seq(0, 1, .1),
       barley_exp = risk * barley_dry + (1 - risk) * barley_wet,
      wheat_exp = risk * wheat_dry + (1 - risk) * wheat_wet) %>%
  gather(type, utility, barley_exp:wheat_exp)
```

@@

```{r, echo = FALSE}
indiff <- tibble(risk = (barley_wet - wheat_wet) / (wheat_dry - wheat_wet - barley_dry + barley_wet),
                 utility = risk * wheat_dry + (1 - risk) * wheat_wet)

plt_dat %>%
  ggplot(aes(risk, utility)) +
  geom_hline(yintercept = indiff$utility, linetype = 2)+
  geom_vline(xintercept = indiff$risk, linetype = 2) +
  geom_line(aes(color = type), size = 1.2) +
  geom_point(data = indiff, color = 'black', size = 3) +
  geom_text(data = indiff, label = '(0.3, 1.1)', nudge_x = 0.1, nudge_y = 0.1) +
  theme_classic()
```

@@

```{r}
# farmer's ne
pmax(0, round((barley_wet - barley_dry) / (wheat_dry + barley_wet - (wheat_wet + barley_dry)), 2))

#nature's ne
round((barley_wet - wheat_wet) / (wheat_dry + barley_wet - (wheat_wet + barley_dry)), 2)
```

@@

And the subjective expected yields for the agents are thus.
```{r}
agents %>%
  mutate(risk = runif(n())) %>%
  mutate(yield_exp = expected_utility(risk),
         barley_exp = risk * barley_dry + (1 - risk) * barley_wet,
         wheat_exp = risk * wheat_dry + (1 - risk) * wheat_wet,
         favorite = case_when(barley_exp == wheat_exp ~ 'either',
                              barley_exp > wheat_exp ~ 'barley',
                              barley_exp < wheat_exp ~ 'wheat')) 
```

@@

Allow the agents to learn the drought risk over time.
```{r}
learn <- function(agents, event) {
  agents %>%
    mutate(alpha = event + alpha,
           beta = beta + 1 - event,
           mu = alpha / (alpha + beta),
           risk = map2_dbl(alpha, beta, ~rbeta(1, .x, .y)))
}
```

@@

```{r}
ggplot(simulation, aes(time, risk)) +
  geom_point(alpha = .1) +
  geom_hline(yintercept = indiff$risk, color = 'red', linetype = 2) +
  theme_classic()
```

@@

```{r}
calc_payoff <- function(x, drought) {
  if (drought == 'TRUE') wheat_yield <- wheat_dry; barley_yield <- barley_dry
  if (drought == 'FALSE') wheat_yield <- wheat_wet; barley_yield <- barley_wet

  x %>%
    mutate(yield = wheat_yield * wheat_prop + barley_yield * (1 - wheat_prop))
}

>>>>

@@

Given these payoffs, what's the point of indifference between the two crop types?
```{r, echo = FALSE}
indiff <- tibble(risk = (barley_wet - wheat_wet) / (wheat_dry - wheat_wet - barley_dry + barley_wet),
                 utility = risk * wheat_dry + (1 - risk) * wheat_wet)

tibble(risk = seq(0, 1, .1),
       barley_exp = risk * barley_dry + (1 - risk) * barley_wet,
      wheat_exp = risk * wheat_dry + (1 - risk) * wheat_wet) %>%
  gather(type, utility, barley_exp:wheat_exp) %>%
  ggplot(aes(risk, utility)) +
  geom_hline(yintercept = indiff$utility, linetype = 2)+
  geom_vline(xintercept = indiff$risk, linetype = 2) +
  geom_line(aes(color = type), size = 1.2) +
  geom_point(data = indiff, color = 'black', size = 3) +
  geom_text(data = indiff, label = '(0.3, 1.1)', nudge_x = 0.1, nudge_y = 0.1) +
  theme_classic()
```

@@

```{r}
# farmer's ne
pmax(0, round((barley_wet - barley_dry) / (wheat_dry + barley_wet - (wheat_wet + barley_dry)), 2))
#nature's ne
round((barley_wet - wheat_wet) / (wheat_dry + barley_wet - (wheat_wet + barley_dry)), 2)
```

@@

@@

Allow the agents to learn the drought risk over time. Also need to calculate the actual payoffs for each agent.

```{r}
learn <- function(agents, drought) {
  if (drought == 'TRUE') wheat_yield <- wheat_dry; barley_yield <- barley_dry
  if (drought == 'FALSE') wheat_yield <- wheat_wet; barley_yield <- barley_wet

  agents %>%
    # farmer's turn
      mutate(yield_exp = expected_utility(risk),
         barley_exp = risk * barley_dry + (1 - risk) * barley_wet,
         wheat_exp = risk * wheat_dry + (1 - risk) * wheat_wet,
         crop = case_when(barley_exp > wheat_exp ~ 'barley',
                              barley_exp < wheat_exp ~ 'wheat',
                              barley_exp == wheat_exp ~ 'either'),
         prediction = map2(alpha, beta, ~rbeta(20, .x, .y)),
         wheat_prop = map_dbl(prediction, ~sum(. < indiff$risk)) / 100) %>%
      select(-prediction) %>%
  # nature's turn
    mutate(payoff = wheat_yield * wheat_prop + barley_yield * (1 - wheat_prop),
           alpha = drought + alpha,
           beta = beta + 1 - drought,
           mu = alpha / (alpha + beta),
           risk = map2_dbl(alpha, beta, ~rbeta(1, .x, .y)))
}
```

@@

```{r, echo = TRUE}
ggplot(simulation, aes(time, risk)) +
  geom_point(alpha = .1) +
  geom_hline(yintercept = indiff$risk, color = 'red', linetype = 2) +
  theme_classic()
```

```{r, echo = TRUE}
simulation %>%
  group_by(time, crop) %>%
  tally() %>%
ggplot(aes(time, n, fill = crop)) +
  geom_bar(width = 1, stat = 'identity') +
  scale_fill_manual(values = c('yellow', 'orange')) +
  theme_classic()
```

These two plots begin to look more alike the more you sample, which makes sense because its an ergodic system.
```{r}
ggplot(simulation, aes(time, wheat_prop)) +
  geom_point(alpha = .3) +
  theme_classic()
```

```{r}
simulation %>%
  group_by(time) %>%
  summarise(famine = sum(payoff == 0)) %>%
ggplot(aes(time, famine)) +
  geom_line()
```

@@
